% CREATED BY DAVID FRISK, 2016
\thispagestyle{plain} % Supress header
\section*{Acknowledgements}
Thank you Sólrún Einarsdóttir for suggesting a fantastic project.
It has been great having you as our supervisor.
You have helped us keep our spirits high.

I would like to thank Nick Smallbone for his time with debugging the usage of QuickSpec, I would not have understood its connection to twee-lib without you.

Further, I would like to thank Mario Carneiro for his introductory course to Lean 4, and for his assistance with introducing Lean's IR to us, this project wouldn't have gotten to where it is without you.

I would also like to thank everyone who helped proofread the paper.
Erik and Adam, you have helped to keep the text understandable, by pointing out more fundamental assumptions of the paper.
Anna, Carl, Samuel, and Simon for reading through the paper and ensuring that there common thread throughout the paper.
William, you reading through the paper while speaking loudly lead to many good catches in our use of the language.

\vspace{0.5cm}
\hfill
Nor Führ, Gothenburg, \today

\vspace{1.5cm}

I will also extend my gratitude to our supervisor, Sólrún.
Especially for being so generous with her time when there were many other important things vying for it.

Also, another thanks to Mario.
Navigating the not so well documented internals of Lean and its code actions would have been a lot tougher without his guidance.

\vspace{0.5cm}
\hfill
Erik Nygren, Gothenburg, \today

\cleardoublepage
